SubsidyExplorer - Step 4: Apply proposal 'definitions' to vessel list

### 2020-05-01 scripts/05_subsidy_reform_definitions.Rmd
final definitions

# Paths to other data files
iuu_dat_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")
costello_2016_path <- here::here("data", "edited", "costello_et_al_2016_upsides_tidy.csv")

@@

# Countries participating in the Northern Agreement for fishing
eu_northern_agreement <- c("NOR", "ISL")
```

@@

In order to determine whether a vessel is engaging in distant water fishing, we need to know something about the sovereign state of the vessel's flag state.

```{r}
# Get pairings of administering and sovereign states from the EEZ data
sovereign_pairings <- temp_vessel_list %>%
  distinct(territory1_iso3, sovereign1_iso3) %>%
  rename(flag = territory1_iso3, flag_sovereign = sovereign1_iso3)
sovereign_pairings$flag_sovereign[sovereign_pairings$flag == "CHN"] <- "CHN"

# Create a manual list of missing sovereign pairings
sovereign_pairings_manual <- tibble(flag = c("HKG", "KHM", "CUW", "BLZ", "CYM", "JOR", "PSE", "JAM", "AIA")) %>%
  mutate(flag_sovereign = case_when(flag == "HKG" ~ "CHN",
                                    flag == "CUW" ~ "NLD",
                                    flag == "CYM" ~ "GBR",
                                    flag == "PSE" ~ "ISR",
                                    flag == "AIA" ~ "GBR",
                                    TRUE ~ flag))

sovereign_pairings <- sovereign_pairings %>%
  bind_rows(sovereign_pairings_manual) %>%
  distinct(flag, flag_sovereign)

# Also fix that on the vessel list
temp_vessel_list$sovereign1_iso3[temp_vessel_list$territory1_iso3 == "CHN"] <- "CHN"

# We then match it back to our vessel list so each entry has an assigned sovereign flag
temp_vessel_list <- temp_vessel_list %>%
  left_join(sovereign_pairings, by = "flag")

# We fist exclude entries where the flag state of the vessel and the administering state of the EEZ in which it is fishing are the same
distant_water_fishing <- temp_vessel_list %>%
  mutate(territory1_match = case_when(!is.na(territory1_iso3) & flag == territory1_iso3 ~ T,
                                      TRUE ~ F),
         territory2_match = case_when(!is.na(territory2_iso3) & flag == territory2_iso3 ~ T,
                                      TRUE ~ F),
         territory3_match = case_when(!is.na(territory3_iso3) & flag == territory3_iso3 ~ T,
                                      TRUE ~ F)) %>%
  dplyr::filter(!territory1_match & !territory2_match & !territory3_match)
```

@@

We do not consider the following to be "intra-EU fishing": 
1. The flag state of the vessel is a member state of the EU, Norway, Svalbard and Jan Mayen, or Iceland and the EEZ in which it is fishing belongs to a territory or distant region of an EU member state (e.g. a French-flagged vessel fishing in the EEZ of French Polynesia).

```{r}
# Exclude intra EU fishing
distant_water_fishing <- distant_water_fishing %>%
  mutate(is_intra_EU = case_when(flag %in% c(eu_countries, eu_northern_agreement, norwegian_territories) & territory1_iso3 %in% c(eu_countries, eu_northern_agreement, norwegian_territories) ~ T,
                                 flag_sovereign %in% c(eu_countries, eu_northern_agreement, norwegian_territories) & territory1_iso3 %in% c(eu_countries, eu_northern_agreement, norwegian_territories) ~ T,
                                 TRUE ~ F)) %>%
  dplyr::filter(!is_intra_EU) %>%
  mutate(distant_water = T) %>%
  dplyr::select(eez_id, ssvid, fao_region, distant_water)
```

### Sovereignity

If the sovereign flag state of a vessel is the same as the sovereign state of the EEZ in which it's fishing, this connection may still be considered to be distant water fishing (e.g. US vessel fishing in Palmyra Atoll), though this is generally not considered to be foreign distant water fishing.

@@

Sovereign fishing is included in our definition of distant water fishing

@@

Using the combined list of IUU vessels (https://iuu-vessels.org/), we want to try to match as many vessels as possible to our vessel list.

@@

### Blank database for IUU vessels
iuu_matching_vessels <- tibble(ssvid = integer(0),
                               flag = character(0),
                               broadcast_shipname = character(0))

### Abbreviated vessel list for matching
vessels_for_matching <- temp_vessel_list %>%
  distinct(ssvid, flag, vessel_class, length_m, tonnage_gt, engine_power_kw, broadcast_shipname, broadcast_callsign, broadcast_imo)

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_ssvid %>% distinct(ssvid, flag, broadcast_shipname))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_callsign %>% distinct(ssvid, flag, broadcast_shipname) %>% dplyr::filter(broadcast_shipname != "ROSE4"))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_imo %>% distinct(ssvid, flag, broadcast_shipname))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(name_matches1 %>% distinct(ssvid, flag, broadcast_shipname)) %>%
  bind_rows(name_matches2 %>% distinct(ssvid, flag, broadcast_shipname))

@@

@@

### OA #1: Overfished as determined by the RAM legacy stock assessment database

@@

costello_2016_dat_clean <- costello_2016_dat %>%
  mutate(region_fao_chr = as.character(region_fao),
         n_regions = nchar(region_fao_chr)/2,
         region_fao_split = str_split(gsub("(.{2})", "\\1 ", region_fao_chr), " ")) %>%
  unnest(cols = c(region_fao_split)) %>%
  mutate(region_fao_split = as.numeric(region_fao_split)) %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(iso3 = case_when(country == "Channel Islands" ~ "JEY",
                          country == "Micronesia" ~ "FSM",
                          country == "Zanzibar" ~ "TZA",
                          TRUE ~ countrycode(country, "country.name", "iso3c")))

# Get EEZ regions for matching
eez_regions <- temp_vessel_list %>%
  distinct(eez_id, fao_region, territory1_iso3, territory2_iso3, territory3_iso3, sovereign1_iso3, sovereign2_iso3, sovereign3_iso3)

# Get RAM stocks only and summarize by iso3 and fao_region
ram_data <- costello_2016_dat_clean %>%
  dplyr::filter(dbase == "RAM") %>%
  group_by(iso3, region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy),
            median_bv_bmsy = median(bv_bmsy)) %>%
  mutate(oa_mean = ifelse(mean_bv_bmsy >= 1, F, T),
         oa_med = ifelse(median_bv_bmsy >= 1, F, T),
         oa1 = ifelse(oa_mean & oa_med, T, F))

ram_by_eez_region <- eez_regions %>%
  left_join(ram_data, by = c("territory1_iso3" = "iso3", "fao_region" = "region_fao_split")) %>%
  arrange(eez_id, fao_region) %>%
  dplyr::select(eez_id, fao_region, oa1)
ram_by_eez_region$oa1[is.na(ram_by_eez_region$oa1)] <- F

# Join back to vessel list
temp_vessel_list <- temp_vessel_list %>%
  left_join(ram_by_eez_region, by = c("eez_id", "fao_region"))
```

### OA #2: Overfished as determined by Costello et al. (2016)

```{r}
# Include all stocks 
all_data <- costello_2016_dat_clean %>%
  group_by(iso3, region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy),
            median_bv_bmsy = median(bv_bmsy)) %>%
  mutate(oa_mean = ifelse(mean_bv_bmsy >= 1, F, T),
         oa_med = ifelse(median_bv_bmsy >= 1, F, T),
         oa2 = ifelse(oa_mean & oa_med, T, F))

all_by_eez_region <- eez_regions %>%
  left_join(all_data, by = c("territory1_iso3" = "iso3", "fao_region" = "region_fao_split")) %>%
  arrange(eez_id, fao_region) %>%
  dplyr::select(eez_id, fao_region, oa2)
all_by_eez_region$oa2[is.na(all_by_eez_region$oa2)] <- F

# Join back to vessel list
temp_vessel_list <- temp_vessel_list %>%
  left_join(all_by_eez_region, by = c("eez_id", "fao_region"))
```

@@

```{r}
# Export final vessel list with definitions
write_csv(temp_vessel_list, paste0(results_dir, "vessel-list-2018-final.csv"))
write_csv(temp_vessel_list, paste0(app_dir, "vessel-list-2018-final.csv"))
```

>>>>

# Paths to other data files
iuu_dat_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")
costello_2016_path <- here::here("data", "edited", "costello_et_al_2016_upsides_tidy.csv")
eez_info_path <- here::here("data", "lookup-tables", "eez_info.csv")

@@

# Countries participating in the Northern Agreement for fishing
eu_countries <- country_lookup$iso3[country_lookup$is_EU]
eu_northern_agreement <- c("NOR", "ISL", "SJM")
```

@@

In order to determine whether a vessel is engaging in distant water fishing, we need to know something about the sovereign state of the vessel's flag state. We have already assigned this.

```{r}
# Get pairings of administering and sovereign states from the vessel data
sovereign_pairings <- temp_vessel_list %>%
  distinct(flag_iso3, flag_sovereign_iso3) %>%
  arrange(flag_iso3)

```{r}
# We fist exclude entries where the flag state of the vessel and the administering state of the EEZ in which it is fishing are the same
distant_water_fishing <- temp_vessel_list %>%
  mutate(territory1_match = case_when(!is.na(eez_territory1_iso3) & flag_iso3 == eez_territory1_iso3 ~ T,
                                      TRUE ~ F),
         territory2_match = case_when(!is.na(eez_territory2_iso3) & flag_iso3 == eez_territory2_iso3 ~ T,
                                      TRUE ~ F),
         territory3_match = case_when(!is.na(eez_territory3_iso3) & flag_iso3 == eez_territory3_iso3 ~ T,
                                      TRUE ~ F)) %>%
  dplyr::filter(!territory1_match & !territory2_match & !territory3_match)
```

@@

We do not consider the following to be "intra-EU fishing" (and is therefore distante water fishing): 
1. The flag state of the vessel is a member state of the EU, Norway, Svalbard and Jan Mayen, or Iceland and the EEZ in which it is fishing belongs to a territory or distant region of an EU member state (e.g. a French-flagged vessel fishing in the EEZ of French Polynesia).

```{r}
# Exclude intra EU fishing
distant_water_fishing <- distant_water_fishing %>%
  mutate(is_intra_EU = case_when(flag_iso3 %in% c(eu_countries, eu_northern_agreement) & eez_territory1_iso3 %in% c(eu_countries, eu_northern_agreement) ~ T,
                                 flag_sovereign_iso3 %in% c(eu_countries, eu_northern_agreement) & eez_territory1_iso3 %in% c(eu_countries, eu_northern_agreement) ~ T,
                                 TRUE ~ F)) %>%
  dplyr::filter(!is_intra_EU) %>%
  mutate(distant_water = T) %>%
  dplyr::select(eez_id, ssvid, fao_region, distant_water)
```

#### A Note on Sovereignity

If the sovereign country of the flag state of a vessel is the same as the sovereign country of the EEZ in which it's fishing, this connection may still be considered to be distant water fishing, though this is generally not considered to be foreign distant water fishing.

@@

Though not foreign, we consider sovereign fishing to still be distant water fishing for the purposes of this analysis.

@@

Using the combined list of IUU vessels (https://iuu-vessels.org/), we match as many vessels as possible to our vessel list.

GFW uses a function called "normalize_shipname" from the `startR` package to normalize vessel names. We apply the same function to the IUU vessel list as one method of facilitating matching.

@@

### Blank database for IUU vessels
iuu_matching_vessels <- tibble(ssvid = integer(0),
                               flag_iso3 = character(0),
                               broadcast_shipname = character(0))

### Abbreviated vessel list for matching
vessels_for_matching <- temp_vessel_list %>%
  distinct(ssvid, flag_iso3, vessel_class, length_m, tonnage_gt, engine_power_kw, broadcast_shipname, broadcast_callsign, broadcast_imo)

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_ssvid %>% distinct(ssvid, flag_iso3, broadcast_shipname))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_callsign %>% distinct(ssvid, flag_iso3, broadcast_shipname) %>% dplyr::filter(broadcast_shipname != "ROSE4"))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(iuu_by_imo %>% distinct(ssvid, flag_iso3, broadcast_shipname))

@@

iuu_matching_vessels <- iuu_matching_vessels %>%
  bind_rows(name_matches1 %>% distinct(ssvid, flag_iso3, broadcast_shipname)) %>%
  bind_rows(name_matches2 %>% distinct(ssvid, flag_iso3, broadcast_shipname))

@@

### Disputed areas

We pull the version of marine regions used by GFW to characterize fishing activity.

```{r}
# EEZ info lookup table from GFW
eez_info <- read_csv(eez_info_path)

# Get disputed regions
disputed_areas <- eez_info %>%
  mutate(is_disputed = case_when(str_detect(reporting_name, "Disputed area") == T ~ T,
                                 TRUE ~ F)) %>%
  dplyr::filter(is_disputed)

# apply to vessel list
temp_vessel_list <- temp_vessel_list %>%
  mutate(is_disputed = ifelse(eez_id %in% disputed_areas$eez_id, T, F))
```

@@

### OA #1/#2: B/Bmsy < 1/0.8 as determined by the RAM legacy stock assessment database

@@

# Separate RAM stocks 
ram_stocks <- costello_2016_dat %>%
  dplyr::filter(dbase == "RAM")

# Get statuses of multinational ram stocks by fao region
multinational_ram_stocks <- ram_stocks %>%
  dplyr::filter(is.na(iso3)) %>%
  group_by(region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy, na.rm = T),
            median_bv_bmsy = mean(bv_bmsy, na.rm = T),
            w_mean_bv_bmsy = weighted.mean(bv_bmsy, w = (msy/n_regions), na.rm = T)) %>%
  mutate(oa1_median = ifelse(median_bv_bmsy >= 1, F, T),
         oa1_w_mean = ifelse(w_mean_bv_bmsy >= 1, F, T),
         oa2_median = ifelse(median_bv_bmsy >= 0.8, F, T),
         oa2_w_mean = ifelse(w_mean_bv_bmsy >= 0.8, F, T)) %>%
  dplyr::select(region_fao_split, oa1_median, oa1_w_mean, oa2_median, oa2_w_mean) %>%
  mutate(eez_id = 0)

# Get statuses of domestic stocks by flag state and fao region
eez_ram_stocks <- ram_stocks %>%
  dplyr::filter(!is.na(iso3))

# Get eez ids from fishing effort
eez_effort <- temp_vessel_list %>%
  distinct(eez_id, fao_region, eez_territory1_iso3, eez_territory2_iso3, eez_territory3_iso3, eez_sovereign1_iso3, eez_sovereign2_iso3, eez_sovereign3_iso3)

# Match
eez_ram_stocks <- eez_ram_stocks %>%
  left_join(eez_effort, by = c("iso3" = "eez_territory1_iso3", "region_fao_split" = "fao_region")) %>%
  group_by(eez_id, region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy, na.rm = T),
            median_bv_bmsy = mean(bv_bmsy, na.rm = T),
            w_mean_bv_bmsy = weighted.mean(bv_bmsy, w = (msy/n_regions), na.rm = T)) %>%
  mutate(oa1_median = ifelse(median_bv_bmsy >= 1, F, T),
         oa1_w_mean = ifelse(w_mean_bv_bmsy >= 1, F, T),
         oa2_median = ifelse(median_bv_bmsy >= 0.8, F, T),
         oa2_w_mean = ifelse(w_mean_bv_bmsy >= 0.8, F, T)) %>%
  dplyr::select(eez_id, region_fao_split, oa1_median, oa1_w_mean, oa2_median, oa2_w_mean)

# Combine 
oa_1_2 <- multinational_ram_stocks %>%
  bind_rows(eez_ram_stocks)

# Join back to vessel list
temp_vessel_list <- temp_vessel_list %>%
  left_join(oa_1_2, by = c("eez_id" = "eez_id", "fao_region" = "region_fao_split"))

temp_vessel_list$oa1_median[is.na(temp_vessel_list$oa1_median)] <- F
temp_vessel_list$oa2_median[is.na(temp_vessel_list$oa2_median)] <- F
temp_vessel_list$oa1_w_mean[is.na(temp_vessel_list$oa1_w_mean)] <- F
temp_vessel_list$oa2_w_mean[is.na(temp_vessel_list$oa2_w_mean)] <- F

### OA #3/4: Overfished as determined by Costello et al. (2016)

```{r}
# Get statuses of multinational ram stocks by fao region
multinational_stocks <- costello_2016_dat %>%
  dplyr::filter(is.na(iso3)) %>%
  group_by(region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy, na.rm = T),
            median_bv_bmsy = mean(bv_bmsy, na.rm = T),
            w_mean_bv_bmsy = weighted.mean(bv_bmsy, w = (msy/n_regions), na.rm = T)) %>%
  mutate(oa3_median = ifelse(median_bv_bmsy >= 1, F, T),
         oa3_w_mean = ifelse(w_mean_bv_bmsy >= 1, F, T),
         oa4_median = ifelse(median_bv_bmsy >= 0.8, F, T),
         oa4_w_mean = ifelse(w_mean_bv_bmsy >= 0.8, F, T)) %>%
  dplyr::select(region_fao_split, oa3_median, oa3_w_mean, oa4_median, oa4_w_mean) %>%
  mutate(eez_id = 0)

# Get statuses of domestic stocks by flag state and fao region
eez_stocks <- costello_2016_dat %>%
  dplyr::filter(!is.na(iso3)) %>%
  left_join(eez_effort, by = c("iso3" = "eez_territory1_iso3", "region_fao_split" = "fao_region")) %>%
  group_by(eez_id, region_fao_split) %>%
  summarize(n_stocks = n_distinct(id_orig),
            mean_bv_bmsy = mean(bv_bmsy, na.rm = T),
            median_bv_bmsy = mean(bv_bmsy, na.rm = T),
            w_mean_bv_bmsy = weighted.mean(bv_bmsy, w = (msy/n_regions), na.rm = T)) %>%
  mutate(oa3_median = ifelse(median_bv_bmsy >= 1, F, T),
         oa3_w_mean = ifelse(w_mean_bv_bmsy >= 1, F, T),
         oa4_median = ifelse(median_bv_bmsy >= 0.8, F, T),
         oa4_w_mean = ifelse(w_mean_bv_bmsy >= 0.8, F, T)) %>%
  dplyr::select(eez_id, region_fao_split, oa3_median, oa3_w_mean, oa4_median, oa4_w_mean)

# Combine 
oa_3_4 <- multinational_stocks %>%
  bind_rows(eez_stocks)

# Join back to vessel list
temp_vessel_list <- temp_vessel_list %>%
  left_join(oa_3_4, by = c("eez_id" = "eez_id", "fao_region" = "region_fao_split"))

temp_vessel_list$oa3_median[is.na(temp_vessel_list$oa3_median)] <- F
temp_vessel_list$oa4_median[is.na(temp_vessel_list$oa4_median)] <- F
temp_vessel_list$oa3_w_mean[is.na(temp_vessel_list$oa3_w_mean)] <- F
temp_vessel_list$oa4_w_mean[is.na(temp_vessel_list$oa4_w_mean)] <- F
```

@@

```{r}
# Export final vessel list with definitions
write_csv(temp_vessel_list, paste0(results_dir, "vessel_list_2018_final.csv"))
write_csv(temp_vessel_list, paste0(app_dir, "vessel_list_2018_final.csv"))
```
